Number of COVID-19 Cases in Canada By Region, Age, and Sex

### 2024-11-23 paper/paper.qmd
Updated model set up of paper

---
title: "Number of COVID-19 Cases Around the World"
subtitle: "Investigating the number of COVID-19 Cases, Tests, and Deaths by Country, Region, Age, and Sex"
author: 
  - Bruce Zhang
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Load the dataset
analysis_data <- read.csv(here("data/02-analysis_data/analysis_data.csv"))

@@

@@

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

@@

@@

# Filter data for the selected 5 countries
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
filtered_data <- analysis_data %>%
  filter(Country %in% selected_countries) %>%
  mutate(
    # Create age intervals
    Age_Interval = cut(Age, breaks = seq(0, 100, by = 10), right = FALSE, include.lowest = TRUE)
  )

@@

# Appendix {-}

>>>>

---
title: "Number of COVID-19 Cases in Canada By Region, Age, and Sex"
subtitle: "Predicting the Factors that Influence the Number of COVID-19 Cases and Deaths"
author: 
  - Bruce Zhang
thanks: "Code and data are available at: [https://github.com/brucejczhang/covid_data](https://github.com/brucejczhang/covid_data)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Load the dataset
analysis_data <- read.csv(here("data/02-analysis_data/analysis_data.csv"))
```

@@

# Load necessary libraries
library(dplyr)
library(knitr)

# Calculate Case Fatality Rate (CFR) and summarize the data
summary_stats <- analysis_data %>%
  mutate(
    CFR = Deaths / (Cases + 1e-6) * 100 # Calculate Case Fatality Rate (percentage)
  ) %>%
  group_by(Region, Sex) %>% # Group by Region and Sex
  summarise(
    Mean_Age = mean(Age, na.rm = TRUE),             # Average age
    Total_Cases = sum(Cases, na.rm = TRUE),         # Total cases
    Total_Deaths = sum(Deaths, na.rm = TRUE),       # Total deaths
    Mean_CFR = mean(CFR, na.rm = TRUE)              # Average Case Fatality Rate
  ) %>%
  ungroup() # Remove grouping for display purposes

# Display the summary table
kable(summary_stats, digits = 2, col.names = c(
  "Region", "Sex", "Mean Age", "Total Cases", "Total Deaths", "Mean CFR (%)"
))

@@

A logistic model has been selected as part of this analysis to examine the relationship between region, sex, age, and the number of cases and deaths of COVID-19.

The logistic regression model predicts the probability of a region or individual being classified as "high risk" based on key predictors. In this case, "high risk" is defined as regions where the death rate (CFR) exceeds the median death rate across the dataset.

The CFR is calculated as: 
\begin{align}
\text{CFR} = \frac{\text{Deaths}}{\text{Cases} + 10^{-6}} \times 100
\end{align}

The \textbf{box-and-whisker plot} visualizes the distribution of predicted probabilities (\( p \)) across different \textbf{age intervals}. Each box represents the interquartile range (IQR) of predicted probabilities, with the line inside the box indicating the median predicted probability for that age interval. Whiskers extend to 1.5 times the IQR, and any points outside this range are considered outliers. This visualization highlights the variability of predicted probabilities within each age group.

The logistic regression model equation is:
\begin{align}
\text{logit}(p) = \ln\left(\frac{p}{1 - p}\right) = \beta_0 + \beta_1 \cdot \text{Tests} + \beta_2 \cdot \text{Age Interval} + \epsilon
\end{align}

where:
\begin{itemize}
    \item $p$: Predicted probability of being high risk ($p = \text{Predicted Probability}$).
    \item $\text{logit}(p)$: Log-odds of the high-risk classification.
    \item $\beta_0$: Intercept of the model.
    \item $\beta_1, \beta_2$: Coefficients for the predictors (e.g., number of tests, age interval).
    \item $\epsilon$: Error term.
\end{itemize}

This model explains how testing rates and age intervals affect the likelihood of a region or demographic being classified as high risk. The box-and-whisker plot shows the variability of predicted probabilities across age groups, helping to identify specific age intervals with consistently higher or lower risks.

@@

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
set.seed(21)

# Step 1: Filter for Urban and Rural Regions Only
filtered_data <- analysis_data %>%
  mutate(
    Death_Rate = Deaths / (Cases + 1e-6), # Calculate death rate (avoid division by zero)
    High_Risk = ifelse(Death_Rate > median(Death_Rate, na.rm = TRUE), 1, 0) # Classify high-risk regions
  )

# Step 2: Logistic Regression Model
# Fit a logistic regression model
logistic_model <- glm(
  High_Risk ~ Tests + Region + Age,
  family = binomial(),
  data = filtered_data
)

# Summary of the model
summary(logistic_model)

# Step 3: Add Predictions to the Dataset
filtered_data <- filtered_data %>%
  mutate(Predicted_Probability = predict(logistic_model, type = "response"))

# Step 4: Visualize the Results
# Plot predicted probabilities by region
ggplot(filtered_data, aes(x = Region, y = Predicted_Probability, fill = Region)) +
  geom_boxplot(alpha = 0.6) +
  labs(
    title = "Predicted Probability of High Risk by Region (Urban vs Rural)",
    x = "Region",
    y = "Predicted Probability of High Risk"
  ) +
  theme_minimal()

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
set.seed(21)

# Step 1: Filter for only 'f' and 'm', and define High Risk
filtered_data <- analysis_data %>%
  mutate(
    Death_Rate = Deaths / (Cases + 1e-6), # Calculate death rate (avoid division by zero)
    High_Risk = ifelse(Death_Rate > median(Death_Rate, na.rm = TRUE), 1, 0) # Classify high-risk regions
  )

# Step 2: Logistic Regression Model
# Fit a logistic regression model
logistic_model <- glm(
  High_Risk ~ Tests + Sex + Age,
  family = binomial(),
  data = filtered_data
)

# Summary of the model
summary(logistic_model)

# Step 3: Add Predictions to the Dataset
filtered_data <- filtered_data %>%
  mutate(Predicted_Probability = predict(logistic_model, type = "response"))

# Step 4: Visualize the Results
# Plot predicted probabilities by sex
ggplot(filtered_data, aes(x = Sex, y = Predicted_Probability, fill = Sex)) +
  geom_boxplot(alpha = 0.6) +
  labs(
    title = "Predicted Probability of High Risk by Sex",
    x = "Sex",
    y = "Predicted Probability of High Risk"
  ) +
  theme_minimal()

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
set.seed(21)

# Step 1: Create Age Intervals and Define High Risk
filtered_data <- analysis_data %>%
  mutate(
    Age_Interval = cut(Age, breaks = seq(0, 100, by = 10), right = FALSE, include.lowest = TRUE), # Create age intervals
    Death_Rate = Deaths / (Cases + 1e-6), # Calculate death rate (avoid division by zero)
    High_Risk = ifelse(Death_Rate > median(Death_Rate, na.rm = TRUE), 1, 0) # Classify high-risk regions
  ) %>%
  filter(!is.na(Age_Interval)) # Remove any rows with missing age intervals

# Step 2: Logistic Regression Model
# Fit a logistic regression model
logistic_model <- glm(
  High_Risk ~ Tests + Age_Interval,
  family = binomial(),
  data = filtered_data
)

# Summary of the model
summary(logistic_model)

# Step 3: Add Predictions to the Dataset
filtered_data <- filtered_data %>%
  mutate(Predicted_Probability = predict(logistic_model, type = "response"))

# Step 4: Visualize the Results
# Plot predicted probabilities by age interval
ggplot(filtered_data, aes(x = Age_Interval, y = Predicted_Probability, fill = Age_Interval)) +
  geom_boxplot(alpha = 0.6) +
  labs(
    title = "Predicted Probability of High Risk by Age Interval",
    x = "Age Interval",
    y = "Predicted Probability of High Risk"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for readability

@@

# Filter data for the selected 5 countries
selected_countries <- c("Canada")
filtered_data <- analysis_data %>%
  filter(Country %in% selected_countries) %>%
  mutate(
    # Create age intervals
    Age_Interval = cut(Age, breaks = seq(0, 100, by = 10), right = FALSE, include.lowest = TRUE)
  )

@@

# Appendix {.unnumbered}
